---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---

## Loading and preprocessing the data

1. Load the data

To load the data, we use the read.csv() command and have a look at the data

```{r}
setwd("C:/Users/Roro7_000/Documents/MOOC/Reproducible research/Project1/RepData_PeerAssessment1/activity")

activity<-read.csv('activity.csv',stringsAsFactors=F)
dfActivity<-as.data.frame(activity)
str(dfActivity)
```


## What is mean total number of steps taken per day?

1.Calculate the total number of steps taken per day

We remove row equals to NA in the data frame thanks to the complete.cases command.

```{r}
dfActivityClean<-dfActivity[complete.cases(dfActivity),]
head(dfActivityClean)
```

We use the plyr packages to sum the number of steps by day.
```{r totalNumberStepswithoutNa}

dfActivityClean$date <- as.Date(format(dfActivityClean$date, format="%Y-%m-%d"))
library(plyr)
totalStepsByDay<-ddply(dfActivityClean, .(date), summarize, total=sum(steps))
totalStepsByDay$total<-as.numeric(totalStepsByDay$total)
totalStepsByDay$date<-as.Date(totalStepsByDay$date)
totalStepsByDay$Source<-'NOT NA'
head(totalStepsByDay)
```

2. Histogram of the total number of steps taken every day

```{r plotTotalNumberStepsWithoutNa}
library(ggplot2)
ggplot(data=totalStepsByDay, aes(x = date, y =total)) + geom_histogram(stat = "identity",, fill="#D55E00",colour="#0072B2") + theme(text = element_text(size=20))+ ggtitle("Number total of steps by day")
```

3.  The mean and median of the total number of steps taken per day

```{r table}
ddply(dfActivityClean, .(date), summarize, total=sum(steps),mean=mean(steps),median=median(steps))
```


## What is the average daily activity pattern?

1. Average number of steps depending on the interval
```{r plotDateIntervel}
averageStepsByinterval<-ddply(dfActivityClean, .(interval), summarize, mean=mean(steps))

maxMeanStep<-max(averageStepsByinterval$mean)
invervalWithMaxStep<-averageStepsByinterval[averageStepsByinterval$mean==maxMeanStep,]$interval

ggplot(data=averageStepsByinterval, aes(x = interval, y =mean)) +  geom_line()+geom_point(data = averageStepsByinterval, aes(x=invervalWithMaxStep, max(mean)), col = 'darkblue',size=5) + geom_text(data = averageStepsByinterval, aes(x=invervalWithMaxStep+250, max(mean), label = "Maximum"),colour='darkblue')
```
 
 2.

The 5-minute interval, on average across all the days in the dataset, which contains the maximum number of steps is the `r invervalWithMaxStep` interval.

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset

```{r nrowDf, echo=FALSE}
nrowBefore<-nrow(dfActivity)
nrowAfter<-nrow(dfActivityClean)
missingValueRow<-nrowBefore-nrowAfter
```

The number of row before remove Na is `r nrowBefore` and after : `r nrowAfter`
So the difference is `r missingValueRow`

2. Strategy to replace missing value

Now, we are going to fill the missing values in the Steps column.
We search which column contains missing values

```{r countNaBycolumn}

na_count <-sapply(dfActivity, function(y) sum(length(which(is.na(y)))))
print(na_count,type="html")

```

With the above table, we can see we only have to fill in the steps column.
Now we use the mean of each day to fill the Nas. That's why we use the averageStepsByinterval data set in the third session. 

We scan all the row of the data set with for loop, we detect the line NA. We fill the value looking for the mean by internal (data set we've created).

```{r newDataFrame}
for(j in 1:nrow(dfActivity)){
  if(is.na(dfActivity[j,]$steps)){
    #we look at the averageStepsByinterval dataframe to match with the coresponding interval value
    dfActivity[j,]$steps<-averageStepsByinterval[averageStepsByinterval$interval==dfActivity[j,]$interval,]$mean
  }
}
```

3. The new data set

```{r}
head(dfActivity)
```


4. Total number of steps taken each day

This is the plot and the mean and median total number of steps taken per day with replacing missing value:

```{r}
totalStepsByDayWithNa<-ddply(dfActivity, .(date), summarize, total=sum(steps))
totalStepsByDayWithNa$total<-as.numeric(totalStepsByDayWithNa$total)
totalStepsByDayWithNa$date<-as.Date(totalStepsByDayWithNa$date)
totalStepsByDayWithNa$Source<-'NA'

ggplot(data=totalStepsByDayWithNa, aes(x = date, y =total))+ geom_histogram(stat = "identity",, fill="#D55E00",colour="#0072B2") + theme(text = element_text(size=20))+ ggtitle("Number total of steps by day")

ddply(dfActivity, .(date), summarize, total=sum(steps),mean=mean(steps),median=median(steps))
```

Do these values differ from the estimates from the first part of the assignment?

We have compared the two histograme (one removing missing value and the other fill it)

```{r}
dataMerged<-rbind(totalStepsByDayWithNa,totalStepsByDay)
ggplot(data=dataMerged, aes(x = date, y =total,fill=Source)) + geom_histogram(stat = "identity",binwidth=1, position="dodge")

```

If we compare this two histograms we can see this is the same, except for the missing value (the red larger rectangles)

And the difference id about the median for each day (0 when we don't replace NA and differente from 0 from some date with the new data set)

## Are there differences in activity patterns between weekdays and weekends?

1. Creation new vector

We use apply to scan all the data set and completed a new column "day" with "weekday" or "weekend"
```{r weekday}

dfActivity[,"day"] <-apply(dfActivity, 1, function(x) if(weekdays(as.Date(x[2])) %in% c("lundi","mardi","mercredi","jeudi","vendredi")){ "weekday"}else{"weekend"})
```

We plot the average number of steps taken per type of day:

```{r}
final<-ddply(dfActivity, .(interval,day), summarize, mean=mean(steps))

ggplot(data=final, aes(x = interval, y =mean)) +  geom_line()  + facet_wrap(~ day)+
    theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, vjust=1))
```
